### 2019-09-30 index.Rmd
New webpage
See index.Rmd

---
title: "Water Utility Rates in California"
#author: "Team BA"
output: 
  html_document: 
    self_contained: no
    keep_md: yes
    theme: cosmo
    toc: TRUE
    toc_float: TRUE
    toc_depth: 2
    code_folding: hide
    highlight: "pygments"
---

Welcome to a clearing house of water data. This page explores water utility prices, consumption, sources, and the number of water quality violations each utility has experienced.

To the left is a table of contents for the following page, but most importantly, immediately below is a map of water prices throughout California for the utilities that have greater than 3,000 customers. And for the data intrepid, at the bottom is the unerlying data consisting of the annual water production and deliveries reports, and rates.

```{r global, include=FALSE}

# Global
viridis_color <- "C"

# File paths
file_parent <- ""
file_paths <- list(
  code = paste0(file_parent, "Code/"),
  data = paste0(file_parent, "Data/"),
  data_owrs = paste0(file_parent, "Data/OWRS/Open-Water-Rate-Specification-master/full_utility_rates/California/")
); rm(file_parent)

###############
# Open Packages
###############
# General packages needed
library(stringr)
library(lubridate)
library(viridis)
# US Census Data
library(tidycensus)
library(maps)
library(sf)
library(leaflet)
library(ggridges)
# Utilities
library(ggfortify)
library(haven)
# Aesthetics
library(hrbrthemes)
library(gcookbook)
library(DT)
# Scraping
library(rvest)
# Analysis
library(glmnet)
library(lfe)
# Laod last to avoid masking functionality
library(tidyverse)

###############
# Open Data
###############
owrs_summary_full <- read_csv(paste0(file_paths$data, "OWRS/summary_table_cleaned.csv")) %>% 
  mutate(tier_number = as.character(tier_number))
ear_deliv <- read_csv(paste0(file_paths$data, "EAR/EAR 2013-2016 DELIVERIES FINAL 06-22-2018.csv"))
ear_prod <- read_csv(paste0(file_paths$data, "EAR/EAR 2013-2016 PRODUCTION FINAL 06-22-2018_1.csv"), 
                     col_types = cols(.default = "c")) %>% 
  bind_rows(read_csv(paste0(file_paths$data, "EAR/EAR 2013-2016 PRODUCTION FINAL 06-22-2018_2.csv"),
                     col_types = cols(.default = "c"))) %>% 
  type_convert()

## Geospatial Data
affected <- st_read(paste0(file_paths$data, "affected"), "affected")
util_sf <- st_read(paste0(file_paths$data, "Water_Districts"), "Water_Districts")

###############
# Clean Data
###############
# Clean the OWRS data
owrs_summary_full %<>% 
  arrange(pwsid) %>% 
  filter(!is.na(Tier_price)) %>% 
  mutate(Tier_volume = if_else(bill_type == "Tiered" & is.na(Tier_volume), Inf, Tier_volume))

# Simplify the OWRS data for the shapefile
owrs_summary <- owrs_summary_full %>%
  group_by(pwsid) %>% 
  mutate(tier_count = n()) %>% 
  ungroup() %>% 
  distinct(utility_name, pwsid, effective_date, bill_frequency, bill_type, 
           service_charge, commodity_charge, bill, tier_count)

# Clean EAR data
## Deliveries - i.e. use cases
ear_deliv <- ear_deliv %>%
  select(PWSID, Water.System.Classification, Year, Month, Date,
         `WATER DELIVERIES TO Single.family.Residential`, `WATER DELIVERIES TO  Multi.family.Residential`,
         `WATER DELIVERIES Total.Delivered Residential IN REVISED UNITS (Total Does not include Landscape Irrigation, Agricultural or to other PWS)`,
         `WATER DELIVERIES TO  Commercial.Institutional`:`WATER DELIVERIES TO  Industrial`,
         `WATER DELIVERIES TO  Agricultural`, `WATER DELIVERIES TO  Landscape.Irrigation`,
         `CALCULATED GPCD (Total delivery to residential in gallons per capita day)`) %>% 
  rename(classification = Water.System.Classification,
         deliv_res = `WATER DELIVERIES TO Single.family.Residential`,
         deliv_res_multi = `WATER DELIVERIES TO  Multi.family.Residential`,
         deliv_com_inst = `WATER DELIVERIES TO  Commercial.Institutional`,
         deliv_ind = `WATER DELIVERIES TO  Industrial`,
         deliv_irig = `WATER DELIVERIES TO  Landscape.Irrigation`,
         deliv_ag = `WATER DELIVERIES TO  Agricultural`,
         deliv_irig = `WATER DELIVERIES TO  Landscape.Irrigation`, 
         deliv_res_no_irig = `WATER DELIVERIES Total.Delivered Residential IN REVISED UNITS (Total Does not include Landscape Irrigation, Agricultural or to other PWS)`,
         deliv_gpcd = `CALCULATED GPCD (Total delivery to residential in gallons per capita day)`) %>% 
  mutate(deliv_res = select(., deliv_res, deliv_res_multi) %>% apply(1, sum, na.rm = T),
         Date = mdy(Date)) %>% 
  select(-deliv_res_multi) %>% 
  replace_na(list(deliv_res = 0, deliv_res_no_irig = 0, deliv_com_inst = 0, deliv_ind = 0,
                  deliv_ag = 0, deliv_irig = 0, deliv_gpcd = 0)) %>% 
  type_convert() %>% 
  mutate(deliv_total = select(., deliv_res, deliv_com_inst, deliv_ag) %>% apply(1, sum, na.rm = T)) %>% 
  mutate_at(vars(deliv_res:deliv_irig), .funs = list(perc = ~ . / deliv_total))

# Production - i.e. sources
ear_prod <- ear_prod %>%
  select(PWSID, Year, Month, Date,
         `WATER PRODUCED FROM GROUNDWATER`:`WATER SOLD TO ANOTHER PUBLIC WATER SYSTEM`,
         `CALCULATED GPCD (Total Potable Produced in gallons per capita day)`) %>% 
  rename(prod_ground = `WATER PRODUCED FROM GROUNDWATER`,
         prod_surface  =`WATER PRODUCED FROM SURFACE WATER`,
         prod_purch = `FINSIHIED WATER PURCHASED OR RECEIVED FROM ANOTHER PUBLIC WATER SYSTEM`,
         prod_sold = `WATER SOLD TO ANOTHER PUBLIC WATER SYSTEM`,
         prod_gpcd = `CALCULATED GPCD (Total Potable Produced in gallons per capita day)`) %>% 
  mutate(Date = mdy(Date)) %>% 
  mutate_at(vars(prod_ground:prod_gpcd), str_replace, "-", as.character(NA)) %>% 
  replace_na(list(prod_ground = 0, prod_surface = 0, prod_purch = 0,
                  prod_sold = 0, prod_gpcd = 0)) %>%
  mutate_at(vars(prod_ground:prod_gpcd), str_remove_all, ",") %>% 
  mutate_at(vars(prod_ground:prod_gpcd), str_remove_all, "\\.00") %>% 
  mutate_at(vars(prod_ground:prod_gpcd), as.numeric) %>% 
  mutate(prod_sold = (-prod_sold),
         prod_total = select(., prod_ground, prod_surface, prod_purch, prod_sold) %>% apply(1, sum, na.rm = T)) %>% 
  mutate_at(vars(prod_ground:prod_sold), .funs = list(perc = ~ . / prod_total))

###############
# Match/Join Data
###############
## Join EAR
ear <- ear_deliv %>% 
  full_join(ear_prod)

# Match Data
wat_ut_xwalk <- util_sf %>%
  as_tibble() %>% 
  select(-geometry) %>% 
  # Clean names to join
  mutate(AGENCYNAME = str_remove_all(AGENCYNAME, "District|Water|Irrigation"),
         AGENCYNAME = str_trim(AGENCYNAME, "both")) %>% 
  # Join
  fuzzyjoin::stringdist_left_join(
    owrs_summary %>% 
      # Clean names to joins
      mutate(utility_name = str_remove_all(utility_name, "District|Water|Irrigation|City Of"),
             utility_name = str_trim(utility_name, "both")) %>% 
      rename(AGENCYNAME = utility_name)
  ) %>% 
  filter(!is.na(pwsid))

# Join the utility data for the visualization
util_sf <- util_sf %>% 
  mutate(AGENCYNAME = str_to_title(AGENCYNAME)) %>% 
  # Add water utility data
  left_join(wat_ut_xwalk) %>% 
  filter(!is.na(pwsid)) %>% 
  # mutate(OBJECTID = as_factor(OBJECTID)) %>% 
  left_join(
    # Add violation counts
    affected %>% 
      select(OBJECTID, Freq) %>% 
      as_tibble() %>% 
      select(-geometry) %>% 
      mutate(OBJECTID = as.numeric(OBJECTID))
  ) %>% 
  mutate_at(vars(commodity_charge, service_charge, bill), round, 2)

# Make a total summary 
ear_summary <- ear %>% 
  mutate(quarter = quarter(Date)) %>% 
  filter(quarter %in% 1:4) %>% 
  group_by(PWSID, quarter) %>% 
  summarise_at(vars(deliv_res:prod_sold_perc), mean, na.rm = T) %>% 
  ungroup() %>% 
  gather(variable, value, -PWSID, -quarter) %>% 
  unite(temp, variable, quarter, sep = "_Q") %>% 
  spread(temp, value) %>% 
  left_join(
    ear %>% 
      group_by(PWSID) %>% 
      summarise_at(vars(deliv_res:prod_sold_perc), mean, na.rm = T) %>% 
      rename_at(vars(deliv_res:prod_sold_perc), funs(paste0(., "_annual")))
  ) %>% 
  rename(pwsid = PWSID) %>% 
  inner_join(wat_ut_xwalk) %>% 
  left_join(
    # Add violation counts
    affected %>% 
      select(OBJECTID, Freq) %>% 
      as_tibble() %>% 
      select(-geometry) %>% 
      mutate(OBJECTID = as.numeric(OBJECTID))
  )

# Add source and sales data to master data
util_sf <- util_sf %>% 
  left_join(
    ear_summary %>% 
      select(pwsid, prod_ground_perc_annual, prod_surface_perc_annual, prod_purch_perc_annual,
             prod_gpcd_annual) %>% 
      mutate_at(vars(prod_ground_perc_annual, prod_surface_perc_annual, prod_purch_perc_annual), ~ (. * 100) %>% round(0)) %>% 
      mutate(prod_gpcd_annual = prod_gpcd_annual %>% round(0))
  )

```

## California's Municipal Water Mapped

*Note*: The map is sparse because it only covers communities with over 3,000 customers.

```{r shiny_app, out.width="100%", echo = FALSE, warning = FALSE, message = FALSE, width = 400, height = 800}

# Creating labels
wat_ut_labels <- sprintf("<strong>%s</strong>
                          <br/>Bill Type: %s
                          <br/>Bill Frequency: %s
                          <br/>Number of Tiers: %s
                          <br/>Service Charge: %s
                          <br/>Consumption Charge: %s
                          <br/>Average Bill: %s
                          <br/>Ave. Consumption (Gal./Day Percapita): %s
                          <br/>Water Quality Violations: %s
                          <br/><strong>Source (Percent): %s</strong>
                          <br/>Groundwater: %s
                          <br/>Surface Water: %s
                          <br/>Purchased: %s",
                  util_sf$AGENCYNAME, util_sf$bill_type, util_sf$bill_frequency,
                  util_sf$tier_count, util_sf$service_charge, util_sf$commodity_charge,
                  util_sf$bill, util_sf$prod_gpcd_annual, util_sf$Freq, "", 
                  util_sf$prod_ground_perc_annual, util_sf$prod_surface_perc_annual, 
                  util_sf$prod_purch_perc_annual) %>%
  lapply(htmltools::HTML)

# Put the service charge in bins
## Service Charges
bins_serv <- seq(min(util_sf$service_charge), max(util_sf$service_charge), length.out = 10)
pal_serv <- colorBin("plasma", util_sf$service_charge, bins = bins_serv)
## Commondity
bins_com <- seq(min(util_sf$commodity_charge), max(util_sf$commodity_charge), length.out = 10)
pal_com <- colorBin("plasma", util_sf$commodity_charge, bins = bins_com)
## Total Bill
bins_bil <- seq(min(util_sf$bill), max(util_sf$bill), length.out = 10)
pal_bil <- colorBin("plasma", util_sf$bill, bins = bins_bil)

pal <- colorNumeric("plasma", domain = util_sf$bill)

# Plot
leaflet(util_sf, width = "100%") %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  setView(lng = -122, lat = 37.9, zoom = 8) %>% 
  addPolygons(color = ~pal_bil(bill),
              fillOpacity = 0.5,
              weight = 1,
              label = wat_ut_labels,
              labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "15px",
                direction = "auto"),
              group = "Average Bill") %>% 
  addPolygons(color = ~pal_serv(service_charge),
              fillOpacity = 0.5,
              weight = 1,
              label = wat_ut_labels,
              labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "15px",
                direction = "auto"),
              group = "Service Charge") %>%
  addPolygons(color = ~pal_com(commodity_charge),
              fillOpacity = 0.5,
              weight = 1,
              label = wat_ut_labels,
              labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "15px",
                direction = "auto"),
              group = "Water Price") %>%
  # addLegend(colors = ~pal_bil(bill),
  #           title = "Average Bill") %>%
  addLayersControl(
    overlayGroups = c("Average Bill", "Service Charge", "Water Price"),
    options = layersControlOptions(collapsed = FALSE),
    position = "bottomright"
  ) %>%
  hideGroup(c("Service Charge", "Water Price")) %>%
  leaflet.extras::addSearchOSM(options = list(position = "topright"))

```

## Water Prices

*Note*: Water price data only covers communities with over 3,000 customers. The average consumer is used to calculate the average service charge (fixed charge paid regardless of your consumption) and consumptive charge (price paid for consumption).

```{r sum_stats, echo = FALSE, warning = FALSE, message = FALSE, fig.height=4.5, fig.width=8, fig.align="center"}

# Distribution of Prices
owrs_summary %>% 
  select(bill, commodity_charge, service_charge) %>% 
  gather() %>% 
  mutate(key = case_when(
    key == "bill" ~ "Average Bill",
    key == "commodity_charge" ~ "Water Price",
    key == "service_charge" ~ "Service Charge",
    TRUE ~ key
  ),
  key = as_factor(key)) %>% 
  ggplot(aes(value, group = key, color= key)) + geom_line(stat = "density") + 
  scale_color_viridis_d(option = viridis_color, end = 0.8) +
  theme_ipsum() +
  labs(color = "Bill Type", x = "Price (USD)", y = "Density") +
  theme(legend.position = "bottom")

# Summary Statistics

owrs_summary_full %>% 
  filter(is.finite(Tier_volume)) %>% 
  # mutate(Tier_volume = if_else(is.infinite(Tier_volume), 100, Tier_volume)) %>% 
  ggplot(aes(x = Tier_volume, y = tier_number, fill = tier_number)) +
  geom_density_ridges(scale = 10, size = 0.25, rel_min_height = 0.03) +
  theme_ridges() +
  scale_x_continuous(limits=c(0, 150), expand = c(0.01, 0)) +
  #scale_y_discrete(expand = c(0.01, 0)) +
  scale_fill_brewer(palette = "YlOrRd")
  #scale_y_reverse(breaks=c(2000, 1980, 1960, 1940, 1920, 1900), expand = c(0.01, 0))

ggplot(owrs_summary_full, aes(x = Tier_price, y = tier_number, fill = tier_number)) +
  geom_density_ridges(scale = 10, size = 0.25, rel_min_height = 0.03) +
  theme_ridges() +
  scale_x_continuous(limits=c(0, 20), expand = c(0.01, 0)) +
  scale_fill_brewer(palette = "YlOrRd")
#scale_y_reverse(breaks=c(2000, 1980, 1960, 1940, 1920, 1900), expand = c(0.01, 0))

owrs_summary_full %>% 
  ggplot(aes(bill, approximate_median_income)) + 
  geom_point(size = 0.6) + 
  theme_ipsum() + 
  stat_smooth(method = "lm") +
  labs(x = "Average Bill (USD)", y = "Median Income (USD)") +
  xlim(c(0, 250))

```

## Extensions and Next Steps

* Defining affordability for water services:
  + A function of household income, number of people in HH, bill, regional average consumption ,and household consumption. Ultimately this is the percent of the income that a household spends on water conditional on regional consumption.
* Underlying details for each utility's produciton and deliveries data.
* Use the rate parser to build a measure of the average bill size.

### Downlaod Underlying Data

```{r data_download, echo=FALSE, warning = FALSE, message = FALSE}

# Names for the data set made public
util_sf_colnames <- c("Agency", "PWSID", "Date Updated", "Bill Frequency", "Bill Type",
                      "Average Fixed Charges", "Average Variable Change", "Average Total Bill",
                      "Number of Tiers at Agency", "Water Quality Violation Count",
                      "Source: Groundwater (%)", "Source: Surface Water (%)", "Source: Purchased (%)",
                      "Average Consumption Percapita per Day (Gallons)")

# Post the data
util_sf %>% 
  as_tibble() %>% 
  select(-geometry, -OBJECTID, -AGENCYUNIQ, -MODIFIEDBY, -SOURCECOMM, -Date_Data_, 
         -GlobalID, -AGENCYNAME.x, -AGENCYNAME.y, -SOURCE, -LASTMODIFI) %>% 
  datatable(., 
            extensions = 'Buttons', 
            colnames = util_sf_colnames,
            options = list(dom = 'Bfrtip', 
                           buttons = c('excel', "csv")))

```